---
title: "First Analysis"
author: "Micaela Chapuis"
date: "2025-02-25"
output: github_document
---

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(seacarb)
library(lubridate)
library(car)
```

## Load in Data
```{r, message = FALSE}
chem <- read_csv(here("Test Sampling", "Data", "tidepool_chemistry.csv"))
params <- read_csv(here("Test Sampling", "Data", "tidepool_parameters.csv"))
pHSlope <- read_csv(here("Test Sampling", "Data", "pHSlope.csv"))
TA <- read_csv(here("Test Sampling", "Data", "total_alkalinity.csv"))
```

## Data cleaning and joining

Make all pool numbers factors
```{r}
chem$pool_number <- factor(chem$pool_number)
params$pool_number <- factor(params$pool_number)
```


Remove junk samples from TA Data
```{r}
junks <- c("junk1", "junk2", "junk3")
TA <- TA %>% filter(!sample_ID %in% junks)
```

Join in TA data
```{r}
data <- left_join(chem, TA, by = "sample_ID")
```


Add in pH calculated from tris and physical parameters
```{r}
data <- left_join(data, pHSlope)
data <- left_join(data, (params %>% select(site, pool_number, substrate, perimeter_m, surface_area_m2)), by = c("site", "pool_number"))
```

Selecting columns 
```{r}
data <- data %>% select(-observers, -notes, Sample.Index)
```

Removing DH Pool 1 (limestone + basalt)
```{r}
data <- data %>% filter(!pool_ID == "dh_p1")
```

Adding in "ocean" as substrate
```{r}
data <- data %>% mutate(substrate = replace_na(substrate, "ocean"))
```

Salinity Normalize TA
```{r}
data <- data %>% mutate(TA_norm = TA*salinity_lab/35)
```


Calculate delta pH and delta TA
```{r}
delta_calc <- data %>% 
  select(date, site, pool_number, substrate, time_point, temp_pool, sample_time, pH, TA_norm)  %>%
  group_by(date, site, pool_number, substrate) %>%  # group by metadata
  arrange(time_point, .by_group = TRUE) %>%  # ensure correct order
  reframe(
    delta_pH = case_when(
      site == "Diamond Head" ~ pH[time_point = 3] - pH[time_point = 1],  # Diamond Head (Time 3 - Time 1)
      site == "Sandy Beach" ~ pH[time_point = 2] - pH[time_point = 1]   # Sandy Beach (Time 2 - Time 1)
    ),
    delta_TA = case_when(
      site == "Diamond Head" ~ TA_norm[time_point = 3] - TA_norm[time_point = 1],  # Diamond Head (Time 3 - Time 1)
      site == "Sandy Beach" ~ TA_norm[time_point = 2] - TA_norm[time_point = 1]   # Sandy Beach (Time 2 - Time 1)
    ), 
    delta_temp = case_when(
      site == "Diamond Head" ~ temp_pool[time_point = 3] - temp_pool[time_point = 1],  # Diamond Head (Time 3 - Time 1)
      site == "Sandy Beach" ~ temp_pool[time_point = 2] - temp_pool[time_point = 1]   # Sandy Beach (Time 2 - Time 1)
    ), 
    delta_time = case_when(
      site == "Diamond Head" ~ as.numeric(difftime(sample_time[time_point = 3], sample_time[time_point = 1], units = "mins")),  # Time 3 - Time 1 for Diamond Head
      site == "Sandy Beach" ~ as.numeric(difftime(sample_time[time_point = 2], sample_time[time_point = 1], units = "mins"))   # Time 2 - Time 1 for Sandy Beach
    )
  ) %>% 
    distinct(date, site, pool_number, substrate, .keep_all = TRUE)  # Ensure only one row per pool
```


Doing it for just two timepoints for now (T1-T2)
```{r} 
delta_calc2 <- data %>% 
  select(date, site, pool_number, substrate, time_point, temp_pool, sample_time, pH, TA_norm)  %>%
  group_by(date, site, pool_number, substrate) %>%  # group by metadata
  arrange(time_point, .by_group = TRUE) %>%  # ensure correct order
  reframe(
    delta_pH = case_when(
      site == "Diamond Head" ~ pH[time_point = 2] - pH[time_point = 1],  # Diamond Head (Time 2 - Time 1)
      site == "Sandy Beach" ~ pH[time_point = 2] - pH[time_point = 1]   # Sandy Beach (Time 2 - Time 1)
    ),
    delta_TA = case_when(
      site == "Diamond Head" ~ TA_norm[time_point = 2] - TA_norm[time_point = 1],  # Diamond Head (Time 2 - Time 1)
      site == "Sandy Beach" ~ TA_norm[time_point = 2] - TA_norm[time_point = 1]   # Sandy Beach (Time 2 - Time 1)
    ), 
    delta_temp = case_when(
      site == "Diamond Head" ~ temp_pool[time_point = 2] - temp_pool[time_point = 1],  # Diamond Head (Time 2 - Time 1)
      site == "Sandy Beach" ~ temp_pool[time_point = 2] - temp_pool[time_point = 1]   # Sandy Beach (Time 2 - Time 1)
    ), 
    delta_time = case_when(
      site == "Diamond Head" ~ as.numeric(difftime(sample_time[time_point = 2], sample_time[time_point = 1], units = "mins")),  # Time 2 - Time 1 for Diamond Head
      site == "Sandy Beach" ~ as.numeric(difftime(sample_time[time_point = 2], sample_time[time_point = 1], units = "mins"))   # Time 2 - Time 1 for Sandy Beach
    )
  ) %>% 
    distinct(date, site, pool_number, substrate, .keep_all = TRUE)  # Ensure only one row per pool
```


Now T2-T3
```{r}
delta_calc3 <- data %>% 
  select(date, site, pool_number, substrate, time_point, temp_pool, sample_time, pH, TA_norm)  %>%
  group_by(date, site, pool_number, substrate) %>%  # group by metadata
  arrange(time_point, .by_group = TRUE) %>%  # ensure correct order
  reframe(
    delta_pH = case_when(
      site == "Diamond Head" ~ pH[time_point = 3] - pH[time_point = 2],  # Diamond Head (Time 3 - Time 2)
      site == "Sandy Beach" ~ pH[time_point = 2] - pH[time_point = 1]   # Sandy Beach (Time 2 - Time 1)
    ),
    delta_TA = case_when(
      site == "Diamond Head" ~ TA_norm[time_point = 3] - TA_norm[time_point = 2],  # Diamond Head (Time 3 - Time 2)
      site == "Sandy Beach" ~ TA_norm[time_point = 2] - TA_norm[time_point = 1]   # Sandy Beach (Time 2 - Time 1)
    ), 
    delta_temp = case_when(
      site == "Diamond Head" ~ temp_pool[time_point = 3] - temp_pool[time_point = 2],  # Diamond Head (Time 3 - Time 2)
      site == "Sandy Beach" ~ temp_pool[time_point = 2] - temp_pool[time_point = 1]   # Sandy Beach (Time 2 - Time 1)
    ), 
    delta_time = case_when(
      site == "Diamond Head" ~ as.numeric(difftime(sample_time[time_point = 3], sample_time[time_point = 2], units = "mins")),  # Time 3 - Time 2 for Diamond Head
      site == "Sandy Beach" ~ as.numeric(difftime(sample_time[time_point = 2], sample_time[time_point = 1], units = "mins"))   # Time 2 - Time 1 for Sandy Beach
    )
  ) %>% 
    distinct(date, site, pool_number, substrate, .keep_all = TRUE)  # Ensure only one row per pool
```


## Data Viz!
### Temperature
```{r}
data %>% ggplot(aes(x = sample_time, y = temp_pool, color = pool_number)) +
  facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "Pool Temperature") + theme(axis.text.x = element_text(angle = 30))
```
```{r}
data %>% filter(site == "Diamond Head") %>%
  ggplot(aes(x = sample_time, y = temp_pool, color = pool_number)) +   facet_wrap(~date, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "Pool Temperature - DH")  + theme(axis.text.x = element_text(angle = 30)) 
```

```{r}
data %>% filter(site == "Sandy Beach") %>%
  ggplot(aes(x = sample_time, y = temp_pool, color = pool_number)) +   facet_wrap(~date, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "Pool Temperature - SB")  + theme(axis.text.x = element_text(angle = 30))
```

### pH
```{r}
data %>% ggplot(aes(x = sample_time, y = pH, color = pool_number)) +
    facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "pH (tris)")  + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% ggplot(aes(x = sample_time, y = pH_probe, color = pool_number)) +
    facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "pH (probe)")  + theme(axis.text.x = element_text(angle = 30))
```
```{r}
data %>% filter(site == "Diamond Head") %>%
  ggplot(aes(x = sample_time, y = pH, color = pool_number)) +   facet_wrap(~date, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "pH (tris) - DH")  + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% filter(site == "Sandy Beach") %>%
  ggplot(aes(x = sample_time, y = pH, color = pool_number)) +   facet_wrap(~date, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "pH (tris) - SB")  + theme(axis.text.x = element_text(angle = 30))
```


### TA
```{r}
data %>% ggplot(aes(x = sample_time, y = TA_norm, color = pool_number)) +   facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "TA normalized")  + theme(axis.text.x = element_text(angle = 30))
```
```{r}
data %>% filter(site == "Diamond Head") %>%
  ggplot(aes(x = sample_time, y = TA_norm, color = pool_number)) +   facet_wrap(~date, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "TA normalized - DH")  + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% filter(site == "Sandy Beach") %>%
  ggplot(aes(x = sample_time, y = TA_norm, color = pool_number)) +   facet_wrap(~date, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "TA normalized - SB")  + theme(axis.text.x = element_text(angle = 30))
```


### Salinity
```{r}
data %>% ggplot(aes(x = sample_time, y = salinity_field, color = pool_number)) +   facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "Salinity (field)")  + theme(axis.text.x = element_text(angle = 30))
```
```{r}
data %>% filter(site == "Diamond Head") %>%
  ggplot(aes(x = sample_time, y = salinity_field, color = pool_number)) +   facet_wrap(~date, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "Salinity (field) - DH")  + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% filter(site == "Sandy Beach") %>%
  ggplot(aes(x = sample_time, y = salinity_field, color = pool_number)) +   facet_wrap(~date, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "Salinity (field) - SB")  + theme(axis.text.x = element_text(angle = 30))
```



### Deltas
#### Delta TA
```{r}
delta_calc %>% ggplot(aes(x = substrate, y = delta_TA, color = substrate)) + #, label = pool_number)) + 
  facet_wrap(~site, scales = "free_x") +
  geom_boxplot(alpha = 0.7) + 
  geom_point() + 
 # geom_text(hjust=0, vjust=0) +
  theme_bw() + 
  scale_color_manual(values = c("gray20", "sienna2", "palegreen3", "dodgerblue3")) +
  guides(color = "none") +
  labs(x = "Substrate", y = "Delta TA", title = "DH T3-T1") + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray")
```

```{r}
delta_calc2 %>% ggplot(aes(x = substrate, y = delta_TA, color = substrate)) + #, label = pool_number)) + 
  facet_wrap(~site, scales = "free_x") +
  geom_boxplot(alpha = 0.7) + 
  geom_point() + 
 # geom_text(hjust=0, vjust=0) +
  theme_bw() + 
  scale_color_manual(values = c("gray20", "sienna2", "palegreen3", "dodgerblue3")) +
  guides(color = "none") +
  labs(x = "Substrate", y = "Delta TA", title = "DH T2-T1") + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray")
```

```{r}
delta_calc3 %>% ggplot(aes(x = substrate, y = delta_TA, color = substrate)) + #, label = pool_number)) + 
  facet_wrap(~site, scales = "free_x") +
  geom_boxplot(alpha = 0.7) + 
  geom_point() + 
 # geom_text(hjust=0, vjust=0) +
  theme_bw() + 
  scale_color_manual(values = c("gray20", "sienna2", "palegreen3", "dodgerblue3")) +
  guides(color = "none") +
  labs(x = "Substrate", y = "Delta TA", title = "DH T3-T2") + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray")
```

#### Delta pH
```{r}
delta_calc %>% ggplot(aes(x = substrate, y = delta_pH, color = substrate)) + #, label = pool_number)) + 
  facet_wrap(~site, scales = "free_x") +
  geom_boxplot(alpha = 0.7) + 
  geom_point() + 
#  geom_text(hjust=0, vjust=0) +
  theme_bw() + 
  scale_color_manual(values = c("gray20", "sienna2", "palegreen3", "dodgerblue3")) +
  guides(color = "none") +
  labs(x = "Substrate", y = "Delta pH", title = "DH T3-T1") + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray")
```

```{r}
delta_calc2 %>% ggplot(aes(x = substrate, y = delta_pH, color = substrate)) + 
  facet_wrap(~site, scales = "free_x") +
  geom_boxplot(alpha = 0.7) + 
  geom_point() + 
  theme_bw() + 
  scale_color_manual(values = c("gray20", "sienna2", "palegreen3", "dodgerblue3")) +
  guides(color = "none") +
  labs(x = "Substrate", y = "Delta pH", title = "DH T2-T1") + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray")
```
```{r}
delta_calc3 %>% ggplot(aes(x = substrate, y = delta_pH, color = substrate)) + 
  facet_wrap(~site, scales = "free_x") +
  geom_boxplot(alpha = 0.7) + 
  geom_point() + 
  theme_bw() + 
  scale_color_manual(values = c("gray20", "sienna2", "palegreen3", "dodgerblue3")) +
  guides(color = "none") +
  labs(x = "Substrate", y = "Delta pH", title = "DH T3-T2") + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray")
```

#### Delta TA vs Delta pH
```{r}
delta_calc %>% 
  filter(substrate == "limestone" | substrate == "basalt") %>%
  ggplot(aes(x = delta_pH, y = delta_TA, color = substrate)) + 
  facet_wrap(~site, scales = "free_x") +
  geom_point() + 
  theme_bw() + 
  scale_color_manual(values = c("gray20", "sienna2")) +
  guides(color = "none") +
  labs(x = "Delta pH", y = "Delta TA", title = "DH T3-T1") 
```

```{r}
delta_calc2 %>% 
  filter(substrate == "limestone" | substrate == "basalt") %>%
  ggplot(aes(x = delta_pH, y = delta_TA, color = substrate)) + 
  facet_wrap(~site, scales = "free_x") +
  geom_point() + 
  theme_bw() + 
  scale_color_manual(values = c("gray20", "sienna2")) +
  guides(color = "none") +
  labs(x = "Delta pH", y = "Delta TA", title = "DH T2-T1") 
```

```{r}
delta_calc3 %>% 
  filter(substrate == "limestone" | substrate == "basalt") %>%
  ggplot(aes(x = delta_pH, y = delta_TA, color = substrate)) + 
  facet_wrap(~site, scales = "free_x") +
  geom_point() + 
  theme_bw() + 
  scale_color_manual(values = c("gray20", "sienna2")) +
  guides(color = "none") +
  labs(x = "Delta pH", y = "Delta TA", title = "DH T3-T2") 
```


#### Delta Temp vs Delta pH
```{r}
delta_calc %>% 
  filter(substrate == "limestone" | substrate == "basalt") %>%
  ggplot(aes(x = delta_temp, y = delta_pH, color = substrate)) + 
  facet_wrap(~site, scales = "free_x") +
  geom_point() + 
  theme_bw() + 
  scale_color_manual(values = c("gray20", "sienna2")) +
  guides(color = "none") +
  labs(x = "Delta Temperature", y = "Delta pH", title = "DH T3-T1") 
```

## Models
### ph vs params
```{r}
model1 <- lm(pH ~ sample_time + perimeter_m + surface_area_m2, data = data)
summary(model1)
Anova(model1)
```

### TA vs params
```{r}
model2 <- lm(TA ~ sample_time + perimeter_m + surface_area_m2, data = data)
summary(model2)
Anova(model2)
```

### ph vs TA
```{r}
model3 <- lm(pH ~ TA, data)
summary(model3)
```

### pH vs substrate
```{r}
model4 <- lm(pH ~ substrate, data)
summary(model4)
Anova(model4)
```

### TA vs substrate
```{r}
model5 <- lm(TA ~ substrate, data)
summary(model5)
Anova(model5)
```

### temp vs substrate
```{r}
model6 <- lm(temp_pool ~ substrate, data)
summary(model6)
Anova(model6)
```


