---
title: "Second Analysis"
author: "Micaela Chapuis"
date: "2025-04-08"
output: github_document
---
## Load Libraries
```{r message = FALSE}
library(tidyverse)
library(here)
library(seacarb)
library(lubridate)
library(car)
library(lubridateExtras)
library(lme4)
library(scales)
library(ggeffects)
library(sjPlot)
```

## Load in Data
```{r, message = FALSE}
chem <- read_csv(here("Test Sampling", "Data", "tidepool_chemistry.csv"))
params <- read_csv(here("Test Sampling", "Data", "tidepool_parameters.csv"))
pHSlope <- read_csv(here("Test Sampling", "Data", "pHSlope.csv"))
TA <- read_csv(here("Test Sampling", "Data", "total_alkalinity.csv"))
```

## Data cleaning and joining

Make all pool numbers factors
```{r}
chem$pool_number <- factor(chem$pool_number)
params$pool_number <- factor(params$pool_number)
```


Remove junk samples from TA Data
```{r}
junks <- c("junk1", "junk2", "junk3")
TA <- TA %>% filter(!sample_ID %in% junks)
```

Join in TA data
```{r}
data <- left_join(chem, TA, by = "sample_ID")
```


Add in pH calculated from tris and physical parameters
```{r}
data <- left_join(data, pHSlope)
data <- left_join(data, (params %>% select(site, pool_number, substrate, perimeter_m, surface_area_m2)), by = c("site", "pool_number"))
```

Selecting columns 
```{r}
data <- data %>% select(-observers, -notes, Sample.Index)
```

Removing DH Pool 1 (limestone + basalt)
```{r}
data <- data %>% filter(!pool_ID == "dh_p1")
```

Adding in "ocean" as substrate
```{r}
data <- data %>% mutate(substrate = replace_na(substrate, "ocean"))
```

Salinity Normalize TA
```{r}
data <- data %>% mutate(TA_norm = TA*salinity_lab/35)
```

Rounding sample time
```{r}
data <- data %>% mutate(round_time = round_hms(sample_time, "hour"),
                        rounded_time = hour(round_time))
```


Calculate delta pH and delta TA
```{r}
delta_calc <- data %>% 
  select(date, site, pool_number, pool_ID, substrate, day_night, time_point, temp_pool, sample_time, pH, TA_norm)  %>%
  group_by(date, site, pool_number, pool_ID, substrate, day_night) %>%  # group by metadata
  arrange(time_point, .by_group = TRUE) %>%  # ensure correct order
  reframe(
    delta_pH = if_else((date == "2025-02-16"), 
                       (pH[time_point = 2] - pH[time_point = 1]), # Sandy Beach 1st time (Time 2 - Time 1)
                       (pH[time_point = 3] - pH[time_point = 1])),   # (Time 3 - Time 1)
    
    delta_TA = if_else((date == "2025-02-16"), 
                       (TA_norm[time_point = 2] - TA_norm[time_point = 1]), # Sandy Beach 1st time (Time 2 - Time 1)
                       (TA_norm[time_point = 3] - TA_norm[time_point = 1])),  # (Time 3 - Time 1)

    delta_temp = if_else((date == "2025-02-16"), 
                       (temp_pool[time_point = 2] - temp_pool[time_point = 1]), # Sandy Beach 1st time (Time 2 - Time 1)
                       (temp_pool[time_point = 3] - temp_pool[time_point = 1])),  # (Time 3 - Time 1)

    delta_time = if_else((date == "2025-02-16"), 
                       (as.numeric(difftime(sample_time[time_point = 2], sample_time[time_point = 1], units = "mins"))), # Sandy Beach 1st time (Time 2 - Time 1)
                       (as.numeric(difftime(sample_time[time_point = 3], sample_time[time_point = 1], units = "mins"))))  # Time 3 - Time 1
    
  ) %>% 
    distinct(date, site, pool_number, substrate, .keep_all = TRUE)  # Ensure only one row per pool
```

Normalize delta pH and TA to by hour
```{r}
delta_calc <- delta_calc %>% mutate(delta_pH_norm = (delta_pH/delta_time)*60,
                                    delta_TA_norm = (delta_TA/delta_time)*60)
```



## Data Viz

```{r}
plot(data$salinity_field, data$salinity_lab)
```

```{r}
data %>% filter(!substrate %in% "ocean") %>% ggplot(aes(x = substrate, y = pH)) + geom_boxplot() + geom_point() + facet_wrap(~rounded_time)
```
```{r}
data %>% filter(!substrate %in% "ocean") %>% ggplot(aes(x = substrate, y = TA_norm)) + geom_boxplot() + geom_point() + facet_wrap(~rounded_time)
```

```{r}
# mean pH for each tidepool
data %>% group_by(site, pool_number) %>% mutate(pH_mean = mean(pH)) %>%
  filter(!substrate %in% "ocean") %>% ggplot(aes(x = substrate, y = pH_mean)) + geom_boxplot() + geom_point() + facet_wrap(~day_night)
```
```{r}
# mean pH for each date
data %>% group_by(date, site) %>% mutate(pH_mean = mean(pH)) %>%
  filter(!substrate %in% "ocean") %>% ggplot(aes(x = substrate, y = pH_mean)) + geom_boxplot() + geom_point() + facet_wrap(~day_night)
# for night the basalt point at the same ph as limestone is DH basalt, for day DH basalt is higher point ~8.4 and ~8.2
```

```{r}
delta_calc %>% filter(!substrate %in% "ocean") %>% ggplot(aes(x = substrate, y = delta_TA)) + geom_boxplot() + geom_point() + facet_wrap(~day_night)
```

```{r}
delta_calc %>% filter(!substrate %in% "ocean") %>% ggplot(aes(x = substrate, y = delta_pH)) + geom_boxplot() + geom_point() + facet_wrap(~day_night)
```

```{r}
data %>% ggplot(aes(x = sample_time, y = temp_pool)) +
    facet_wrap(~substrate) + geom_smooth() + geom_point() + theme_minimal() + labs(title = "Pool Temp")  + theme(axis.text.x = element_text(angle = 30))
```
```{r}
# committee meeting fig
data %>% ggplot(aes(x = sample_time, y = temp_pool, color = substrate)) +
    facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone", ocean = "Ocean"))) + 
  geom_smooth(method = "gam", se=FALSE) + 
  geom_point() + theme_bw() +
  labs(x = "Sample Time", y = "Temperature (°C)")  + 
  guides(color = "none") +
  theme(axis.text.x = element_text(angle = 25)) 
```



```{r}
data %>% ggplot(aes(x = sample_time, y = salinity_field)) +
    facet_wrap(~substrate) + geom_smooth() + geom_point() + theme_minimal() + labs(title = "Salinity (Field)")  + theme(axis.text.x = element_text(angle = 30))
```
```{r}
data %>% ggplot(aes(x = sample_time, y = salinity_field, color= site)) +
    facet_wrap(~substrate) + geom_smooth() + geom_point() + theme_minimal() + labs(title = "Salinity (Field)")  + theme(axis.text.x = element_text(angle = 30))
```
```{r}
# committee meeting fig
data %>% ggplot(aes(x = sample_time, y = salinity_field, color = substrate)) +
    facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone", ocean = "Ocean"))) + 
  geom_smooth(method = "gam", se=FALSE) + 
  geom_point() + theme_bw() +
  labs(x = "Sample Time", y = "Salinity")  + 
  guides(color = "none") +
  theme(axis.text.x = element_text(angle = 25)) 
```




```{r}
data %>% ggplot(aes(x = sample_time, y = pH, color = substrate)) +
    facet_wrap(~substrate) + 
  geom_smooth(method = "gam", se=FALSE) + 
  geom_point() + theme_minimal() +
  labs(title = "pH (tris)")  + 
  theme(axis.text.x = element_text(angle = 30))
```

```{r}
# committee meeting fig
data %>% ggplot(aes(x = sample_time, y = pH, color = substrate)) +
    facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone", ocean = "Ocean"))) + 
  geom_smooth(method = "gam", se=FALSE) + 
  geom_point() + theme_bw() +
  labs(x = "Sample Time")  + 
  guides(color = "none") +
  theme(axis.text.x = element_text(angle = 25)) 
```

```{r}
data %>% ggplot(aes(x = sample_time, y = TA_norm)) +
    facet_wrap(~substrate) +
  geom_smooth(method = "gam", se=FALSE) + 
  geom_point() + 
  theme_minimal() + 
  labs(title = "TA Norm")  +
  theme(axis.text.x = element_text(angle = 30))
```

```{r}
# committee meeting fig
data %>% ggplot(aes(x = sample_time, y = TA_norm, color = site)) +
    facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone", ocean = "Ocean"))) + 
  geom_smooth(method = "gam", se=FALSE) + 
  geom_point() + theme_bw() +
  labs(x = "Sample Time", y = "Total Alkalinity")  + 
  guides(color = "none") +
  theme(axis.text.x = element_text(angle = 25)) 
```


```{r}
# committee meeting figure
data %>% 
  filter(!substrate %in% "ocean") %>% 
  ggplot(aes(x = pH, y = TA_norm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone"))) +
  theme_bw() + 
  labs(x = "pH", y = "Total Alkalinity") 
```

```{r}
# committee meeting figure
delta_calc %>% 
  filter(substrate == "limestone" | substrate == "basalt") %>%
  ggplot(aes(x = delta_pH_norm, y = delta_TA_norm)) + 
  facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone", ocean = "Ocean"))) +
  geom_point(aes(color = date)) + 
  geom_smooth(method = "lm") + 
  theme_bw() + 
  labs(x = "Delta pH/Hour", y = "Delta TA/Hour") 
```
```{r}

delta_calc %>% 
  filter(substrate == "limestone" | substrate == "basalt") %>%
  ggplot(aes(x = delta_pH_norm, y = delta_temp)) + 
  facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone", ocean = "Ocean"))) +
  geom_point(aes(color = date)) + 
  geom_smooth(method = "lm") + 
  theme_bw() + 
  labs(x = "Delta pH/Hour", y = "Delta Temp") 
```



-----
```{r message = FALSE}
#producers <- read_csv(here("Test Sampling", "Data", "prodcover.csv"))
producers <- read_csv(here("Test Sampling", "Data", "producers.csv"))
new_data <- left_join(data, producers, by = "pool_ID")
new_delta_calc <- left_join(delta_calc, producers, by = "pool_ID")
```


```{r}
new_delta_calc %>% ggplot(aes(x = (turf_cyano_cover+macro_cover), y = delta_pH)) + geom_point()
```


Mean Ph

```{r}
mean_ph <- data %>% group_by(pool_ID, substrate) %>% summarise(mean_pH = mean(pH)) 
mean_pH_day <- data %>% filter(day_night == "day") %>% group_by(pool_ID, substrate) %>% summarise(mean_pH = mean(pH))

mean_ph <- left_join(mean_ph, producers, by = "pool_ID")
mean_pH_day <- left_join(mean_pH_day, producers, by = "pool_ID")
```

```{r}
mod1 <- lm(mean_pH ~ turf_cyano_cover*substrate.x + macro_cover*substrate.x, data = mean_ph)
summary(mod1)
anova(mod1)
plot(ggeffect(mod1))
plot_model(mod1, type = "eff", terms = c("turf_cyano_cover", "substrate.x"))
plot_model(mod1, type = "eff", terms = c("macro_cover", "substrate.x"))
```


```{r}
# committee meeting figure
mean_ph %>% filter(!substrate.x == "ocean") %>%
  ggplot(aes(x = (turf_cyano_cover+macro_cover+CCA_cover), y = mean_pH, color = substrate.x)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x = "Producer Percent Cover", y = "Mean pH", color = "Substrate") +
  theme_bw() +
  scale_color_manual(values = c('#e6194b','#3cb44b'), labels = c("Basalt", "Limestone"))
```


```{r}
#mean_pH_day %>% ggplot(aes(x = (turf_cyano_cover+macro_cover), y = mean_pH, color = substrate.x)) + 
#  geom_point() +
#  geom_smooth(method = "lm", se = FALSE)

mean_pH_day %>% filter(!substrate.x == "ocean") %>%
  ggplot(aes(x = (turf_cyano_cover+macro_cover+CCA_cover), y = mean_pH, color = substrate.x)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x = "Producer Percent Cover", y = "Mean pH", color = "Substrate") +
  theme_bw() +
  scale_color_manual(values = c('#e6194b','#3cb44b'), labels = c("Basalt", "Limestone"))

```

