---
title: "TP_analysis"
author: "Micaela Chapuis"
date: "2025-06-30"
output: github_document
---

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(seacarb)
library(lubridate)
library(car)
library(lubridateExtras)
library(lme4)
library(scales)
library(ggeffects)
library(sjPlot)
```

## Load in Data
```{r, message = FALSE}
chem <- read_csv(here("Data", "tidepool_chemistry.csv"))
params <- read_csv(here("Data", "tidepool_parameters.csv"))
pHSlope <- read_csv(here("Data", "pHSlope.csv"))
TA <- read_csv(here("Data", "total_alkalinity.csv"))
```


## Data cleaning and joining

Make all pool numbers factors
```{r}
chem$pool_number <- factor(chem$pool_number, levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30, "ocean"))
params$pool_number <- factor(params$pool_number)
```


Remove junk samples from TA Data
```{r}
junks <- c("junk1", "junk2", "junk3")
TA <- TA %>% filter(!sample_ID %in% junks)
```

Join in TA data
```{r}
data <- left_join(chem, TA, by = "sample_ID")
```


Add in pH calculated from tris and physical parameters
```{r}
data <- left_join(data, pHSlope)
data <- left_join(data, (params %>% select(site, pool_number, substrate, perimeter_m, surface_area_m2)), by = c("site", "pool_number"))
```

Selecting columns 
```{r}
data <- data %>% select(-observers, -notes, -Sample.Index)
```

Adding in "ocean" as substrate
```{r}
data <- data %>% mutate(substrate = replace_na(substrate, "ocean"))
```

Salinity Normalize TA
```{r}
data <- data %>% mutate(TA_norm = TA*salinity_lab/35)
```


Calculate delta pH and delta TA
```{r}
delta_calc <- data %>% 
  select(date, site, pool_number, substrate, day_night, time_point, temp_pool, sample_time, pH, TA_norm)  %>%
  group_by(date, site, pool_number, substrate) %>%  # group by metadata
  arrange(time_point, .by_group = TRUE) %>%  # ensure correct order
  reframe(
    delta_pH = pH[time_point = 3] - pH[time_point = 1],   # (Time 3 - Time 1)
    
    delta_TA = (TA_norm[time_point = 3] - TA_norm[time_point = 1]),  # (Time 3 - Time 1)

    delta_temp = (temp_pool[time_point = 3] - temp_pool[time_point = 1]),  # (Time 3 - Time 1)

    delta_time = (as.numeric(difftime(sample_time[time_point = 3], sample_time[time_point = 1], units = "mins")))  # Time 3 - Time 1
    
  ) %>% 
    distinct(date, site, pool_number, substrate, .keep_all = TRUE)  # Ensure only one row per pool
```

Normalize delta pH and TA to by hour
```{r}
delta_calc <- delta_calc %>% mutate(delta_pH_norm = (delta_pH/delta_time)*60,
                                    delta_TA_norm = (delta_TA/delta_time)*60)
```


## Data Viz
```{r}
# checking field salinity (with Orion) to lab salinity (with Orion) measurements
plot(data$salinity_field, data$salinity_lab)
```

### Pool Temperature
```{r}
data %>% filter(site %in% "Kaihalulu Beach") %>% ggplot(aes(x = sample_time, y = temp_pool, color = pool_number)) +
  facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "Pool Temperature") + theme(axis.text.x = element_text(angle = 30))
```
```{r}
data %>% filter(site %in% "Sandy Beach") %>% ggplot(aes(x = sample_time, y = temp_pool, color = pool_number)) +
  facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "Pool Temperature") + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% ggplot(aes(x = sample_time, y = temp_pool)) +
    facet_wrap(~substrate) + geom_smooth() + geom_point() + theme_minimal() + labs(title = "Pool Temp")  + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% ggplot(aes(x = sample_time, y = temp_pool, color = substrate)) +
    facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone", ocean = "Ocean"))) + 
  geom_smooth(method = "gam", se=FALSE) + 
  geom_point() + theme_bw() +
  labs(x = "Sample Time", y = "Temperature (°C)")  + 
  guides(color = "none") +
  theme(axis.text.x = element_text(angle = 25)) 
```


### Pool Salinity
```{r}
data %>% filter(site %in% "Kaihalulu Beach") %>% ggplot(aes(x = sample_time, y = salinity_pool, color = pool_number)) +   facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "Salinity (pool)")  + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% filter(site %in% "Sandy Beach") %>% ggplot(aes(x = sample_time, y = salinity_pool, color = pool_number)) +   facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "Salinity (pool)")  + theme(axis.text.x = element_text(angle = 30))
```


```{r}
data %>% ggplot(aes(x = sample_time, y = salinity_pool)) +
    facet_wrap(~substrate) + geom_smooth() + geom_point() + theme_minimal() + labs(title = "Salinity (Pool)")  + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% ggplot(aes(x = sample_time, y = salinity_pool, color = substrate)) +
    facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone", ocean = "Ocean"))) + 
  geom_smooth(method = "gam", se=FALSE) + 
  geom_point() + theme_bw() +
  labs(x = "Sample Time", y = "Pool Salinity")  + 
  guides(color = "none") +
  theme(axis.text.x = element_text(angle = 25)) 
```

### pH
```{r}
data %>% filter(site %in% "Kaihalulu Beach") %>% ggplot(aes(x = sample_time, y = pH, color = pool_number)) +
    facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "pH (tris)")  + theme(axis.text.x = element_text(angle = 30))
```
```{r}
data %>% filter(site %in% "Sandy Beach") %>% ggplot(aes(x = sample_time, y = pH, color = pool_number)) +
    facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "pH (tris)")  + theme(axis.text.x = element_text(angle = 30))
```


```{r}
data %>% filter(site %in% "Kaihalulu Beach") %>% ggplot(aes(x = sample_time, y = pH_probe, color = pool_number)) +
    facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "pH (probe)")  + theme(axis.text.x = element_text(angle = 30))
```
```{r}
data %>% filter(site %in% "Sandy Beach") %>% ggplot(aes(x = sample_time, y = pH_probe, color = pool_number)) +
    facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "pH (probe)")  + theme(axis.text.x = element_text(angle = 30))
```
```{r}
data %>% ggplot(aes(x = sample_time, y = pH)) +
    facet_wrap(~substrate) + geom_smooth() + geom_point() + theme_minimal() + labs(title = "pH")  + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% ggplot(aes(x = sample_time, y = pH, color = substrate)) +
    facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone", ocean = "Ocean"))) + 
  geom_smooth(method = "lm", se=FALSE) + 
  geom_point() + theme_bw() +
  labs(x = "Sample Time")  + 
  guides(color = "none") +
  theme(axis.text.x = element_text(angle = 25)) 
```


### TA
```{r}
data %>% filter(site %in% "Kaihalulu Beach") %>% ggplot(aes(x = sample_time, y = TA_norm, color = pool_number)) +   facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "TA normalized")  + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% filter(site %in% "Sandy Beach") %>% ggplot(aes(x = sample_time, y = TA_norm, color = pool_number)) +   facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "TA normalized")  + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% ggplot(aes(x = sample_time, y = TA_norm)) +
    facet_wrap(~substrate) +
  geom_smooth(method = "lm", se=FALSE) + 
  geom_point() + 
  theme_minimal() + 
  labs(title = "TA Norm")  +
  theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% ggplot(aes(x = sample_time, y = TA_norm, color = substrate)) +
    facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone", ocean = "Ocean"))) + 
  geom_smooth(method = "lm", se=FALSE) + 
  geom_point() + theme_bw() +
  labs(x = "Sample Time", y = "Total Alkalinity")  + 
  guides(color = "none") +
  theme(axis.text.x = element_text(angle = 25)) 
```

### O2
```{r}
data %>% filter(site %in% "Kaihalulu Beach") %>% ggplot(aes(x = sample_time, y = dissolved_oxygen, color = pool_number)) + facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "DO")  + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% filter(site %in% "Sandy Beach") %>% ggplot(aes(x = sample_time, y = dissolved_oxygen, color = pool_number)) +   facet_wrap(~site, scales = "free_x") + geom_line(linewidth = 0.8) + geom_point() + theme_minimal() + labs(title = "DO")  + theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% ggplot(aes(x = sample_time, y = dissolved_oxygen)) +
    facet_wrap(~substrate) +
  geom_smooth(method = "lm", se=FALSE) + 
  geom_point() + 
  theme_minimal() + 
  labs(title = "DO")  +
  theme(axis.text.x = element_text(angle = 30))
```

```{r}
data %>% ggplot(aes(x = sample_time, y = dissolved_oxygen, color = substrate)) +
    facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone", ocean = "Ocean"))) + 
  geom_smooth(method = "lm", se=FALSE) + 
  geom_point() + theme_bw() +
  labs(x = "Sample Time", y = "Dissolved Oxygen")  + 
  guides(color = "none") +
  theme(axis.text.x = element_text(angle = 25)) 
```


### Deltas
```{r}
delta_calc %>% ggplot(aes(x = substrate, y = delta_pH, color = substrate)) + #, label = pool_number)) + 
  geom_boxplot(alpha = 0.7) + 
  geom_point() + 
 # geom_text(hjust=0, vjust=0) +
  theme_bw() + 
  scale_color_manual(values = c("gray20", "sienna2", "palegreen3", "dodgerblue3")) +
  guides(color = "none") +
  labs(x = "Substrate", y = "Delta pH") + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray")
```

```{r}
delta_calc %>% ggplot(aes(x = substrate, y = delta_TA, color = substrate)) + #, label = pool_number)) + 
  geom_boxplot(alpha = 0.7) + 
  geom_point() + 
 # geom_text(hjust=0, vjust=0) +
  theme_bw() + 
  scale_color_manual(values = c("gray20", "sienna2", "palegreen3", "dodgerblue3")) +
  guides(color = "none") +
  labs(x = "Substrate", y = "Delta TA") + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray")
```

### Delta TA vs Delta pH
```{r}
# committee meeting figure
data %>% 
  filter(!substrate %in% "ocean") %>% 
  ggplot(aes(x = pH, y = TA_norm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone"))) +
  theme_bw() + 
  labs(x = "pH", y = "Total Alkalinity") 
```
```{r}
# committee meeting figure
delta_calc %>% 
  filter(substrate == "limestone" | substrate == "basalt") %>%
  ggplot(aes(x = delta_pH_norm, y = delta_TA_norm)) + 
  facet_wrap(~substrate, 
               labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone", ocean = "Ocean"))) +
  geom_point() + 
  geom_smooth(method = "lm") + 
  theme_bw() + 
  labs(x = "Delta pH/Hour", y = "Delta TA/Hour") 
```