---
title: "TP_community_composition"
author: "Micaela Chapuis"
date: "2025-09-01"
output: github_document
---

## Load Libraries
```{r message = FALSE}
library(tidyverse)
library(here)
```

## Load in Data
```{r, message = FALSE}
benthic <- read_csv(here("Data", "benthic_community_composition.csv"))
mobile <- read_csv(here("Data", "mobile_community_composition.csv"))
tp_parameters <- read_csv(here("Data", "tidepool_parameters.csv"))
categories <- read_csv(here("Data", "community_composition_categories.csv"))
```

Plot number of points per pool over time just to check (should be mostly straight lines)
```{r}
benthic %>% filter(site %in% "Kaihalulu Beach") %>% ggplot(aes(x = date, y = total_points, color = factor(pool_number))) + geom_point() + geom_line() 
```

```{r}
benthic %>% filter(site %in% "Sandy Beach") %>% ggplot(aes(x = date, y = total_points, color = factor(pool_number))) + geom_point() + geom_line() 
```


## Benthic

Make a column for month
```{r}
benthic$month <- month(benthic$date, label = TRUE)
```


```{r}
benthic_long <- benthic %>%
  select(-c("notes":"observers")) %>%
  pivot_longer(
    cols = c("bare_rock":"red_sponge"),
    names_to = "id",
    values_to = "num_points")
```

Assign species to broader categories by joining with other datasheet
```{r}
benthic_categories <- left_join(benthic_long, categories)
```


Calculate percent cover by species
```{r}
benthic_long <- benthic_long %>%
                mutate(spp_percent_cover = (num_points/total_points)*100)
```
Plot
```{r}
benthic_long %>% ggplot(aes(x= factor(pool_number),
               y = num_points, 
               fill= factor(id), 
               color= id)) + # set lines surrounding each color to match the fill colors 
      geom_bar(stat="identity", position="fill") + # stacked bars
      facet_wrap(~substrate, scales = "free_x",
                 labeller = as_labeller(c(basalt = "Basalt", limestone = "Limestone"))) + 
  
      labs(x = "Pool", # labels
           y="Relative Abundance",
           fill = "Category") +

      theme_minimal() +  # theme
      theme(title = element_text(size = 18, face = "bold"), # make title bigger and bold
            axis.title = element_text(size = 16), # make all text bigger
            axis.text = element_text(size = 14),
            legend.title = element_text(size = 16),
            legend.text = element_text(size = 14)) +
    
      
      guides(color = "none") #+ # keep only legend for fill since fill and color are the same
      
#      scale_fill_manual(values = c('#e6194b', '#3cb44b', "lightblue", '#4363d8', '#f58231', '#000075', '#46f0f0', '#f032e6', '#fabebe', '#800000', '#008080',  '#e6beff', '#aaffc3', '#ffe119' , '#808000')) +
      
#      scale_color_manual(values = c('#e6194b', '#3cb44b', "lightblue", '#4363d8', '#f58231', '#000075', '#46f0f0', '#f032e6', '#fabebe', '#800000', '#008080',  '#e6beff', '#aaffc3', '#ffe119' , '#808000'))
```


Calculate percent cover by category
```{r}
benthic_categories_pcover <- benthic_categories %>%
                             mutate(cat_percent_cover = (num_points/total_points)*100) %>%
                             select(-c(total_points, id, num_points, production, calcification)) %>%
                             group_by(date, site, water_date, pool_number, substrate, category, category_name) %>%
                             mutate(cat_percent_cover = sum(cat_percent_cover)) %>%
                             distinct(date, site, water_date, pool_number, substrate, .keep_all = TRUE)  # Ensure only one row per pool
```


Change factor order 
```{r}
# Change order levels 
benthic_categories$category_name <- factor(benthic_categories$category_name, 
                                             levels=c("Bare Rock", "Loose Rocks/Rubble", "Sand", "Turf/Cyanobacteria", "CCA", "Encrusting Algae", "Brown Algae", "Red Algae", "Green Algae", "Sponge"))
```


```{r}
benthic_categories %>% 
  filter(!pool_number %in% c(27, 30)) %>%
  ggplot(aes(x= factor(pool_number),
               y = num_points, 
               fill= factor(category_name), 
               color= category_name)) + # set lines surrounding each color to match the fill colors
      geom_bar(stat="identity", position="fill") + # stacked bars
      facet_grid(month ~substrate, scales = "free_x") +
  
      labs(x = "Pool", # labels
           y="Relative Abundance",
           fill = "Category") +

      theme_minimal() +  # theme
      theme(title = element_text(size = 18, face = "bold"), # make title bigger and bold
            axis.title = element_text(size = 16), # make all text bigger
            axis.text = element_text(size = 14),
            legend.title = element_text(size = 16),
            legend.text = element_text(size = 14)) +
    
      
      guides(color = "none")  + # keep only legend for fill since fill and color are the same
      
      scale_fill_manual(values = c('gray75', 'gray30', '#ffe119', '#4363d8', "hotpink", '#aaffc3', '#f58231', '#e6194b', '#3cb44b', "#7C4585")) +
      
      scale_color_manual(values = c('gray75', 'gray30', '#ffe119', '#4363d8', "hotpink", '#aaffc3', '#f58231', '#e6194b', '#3cb44b', "#7C4585"))
```

Benthic Producers
Calculate percent cover for producer/non-producers
```{r}
benthic_producers <- benthic_categories %>%
                             mutate(cat_percent_cover = (num_points/total_points)*100) %>%
                             select(-c(total_points, id, num_points, category, category_name, calcification)) %>%
                             group_by(date, site, water_date, pool_number, substrate, production) %>%
                             mutate(prod_percent_cover = sum(cat_percent_cover)) %>%
                             distinct(date, site, water_date, pool_number, substrate, .keep_all = TRUE)  # Ensure only one row per pool
```


```{r}
benthic_categories %>% 
  filter(!pool_number %in% c(27, 30)) %>%
  ggplot(aes(x= factor(pool_number),
               y = num_points, 
               fill= factor(production), 
               color= production)) + # set lines surrounding each color to match the fill colors
      geom_bar(stat="identity", position="fill") + # stacked bars
      facet_grid(month ~substrate, scales = "free_x") +
  
      labs(x = "Pool", # labels
           y="Relative Abundance",
           fill = "Category")  +
  
      guides(color = "none")  + # keep only legend for fill since fill and color are the same
      theme_minimal() 
```

Benthic Calcifiers
```{r}
benthic_calcifiers <- benthic_categories %>%
                             mutate(cat_percent_cover = (num_points/total_points)*100) %>%
                             select(-c(total_points, id, num_points, category, category_name, production)) %>%
                             group_by(date, site, water_date, pool_number, substrate, calcification) %>%
                             mutate(calc_percent_cover = sum(cat_percent_cover)) %>%
                             distinct(date, site, water_date, pool_number, substrate, .keep_all = TRUE)  # Ensure only one row per pool
```


```{r}
benthic_categories %>% 
  filter(!pool_number %in% c(27, 30)) %>%
  ggplot(aes(x= factor(pool_number),
               y = num_points, 
               fill= factor(calcification), 
               color= calcification)) + # set lines surrounding each color to match the fill colors
      geom_bar(stat="identity", position="fill") + # stacked bars
      facet_grid(month ~substrate, scales = "free_x") +
  
      labs(x = "Pool", # labels
           y="Relative Abundance",
           fill = "Category")  +
  
      guides(color = "none")  + # keep only legend for fill since fill and color are the same
      theme_minimal() 
```


## Mobile
Join with tidepool parameters to get surface area of each pool
```{r}
mobile <- left_join(mobile, tp_parameters, by = c("pool_number", "site", "substrate"))
```

Make a column for month
```{r}
mobile$month <- month(mobile$date, label = TRUE)
```

```{r}
mobile_long <- mobile %>%
  select(-c("notes.x":"perimeter_m", "depth":"notes.y")) %>%
  pivot_longer(
    cols = c("hermit_crab_large":"shrimp"),
    names_to = "id",
    values_to = "count")
```

Assign species to broader categories by joining with other datasheet
```{r}
mobile_categories <- left_join(mobile_long, categories)
```

Calculate density
```{r}
mobile_categories_density <- mobile_categories %>%
                             mutate(density = (count/surface_area_m2)) %>%
                             select(-c(count, id, surface_area_m2)) %>%
                             group_by(date, site, water_date, pool_number, substrate, category, category_name) %>%
                             mutate(density = sum(density)) %>%
                             distinct(date, site, water_date, pool_number, substrate, .keep_all = TRUE)  # Ensure only one row per pool
```

```{r}
mobile_categories_density %>% 
  filter(!pool_number %in% c(27,30)) %>%
  ggplot(aes(x = factor(pool_number),
             y = log(density), 
             fill = category, 
             color = category)) +
  geom_bar(stat = "identity") +
  facet_grid(month~substrate, scale = "free_x") 
```

